---
output: html_document
---

The method below utilizes the ideas from: “Discovering latent association structure via Bayesian one-mode projection of temporal bipartite graphs.”  
<br>
The first level of the model generates a latent association graph, G, on the items.  
<br>
A related problem is the one-mode projection of a bipartite network. In this operation, a bipartite network $G=\{U,V,E\}$, described by the $N\times K$ incidence matrix B is mapped to a graph with only one class of nodes, ${{G}_{U}}=\{V,{{E}_{U}}\}$ via $B:N\times K\to W:N\times N$. The new connections are placed between the nodes of a set U based on the way they linked to the “vanished” set V. In our case, the set U consists of the items and the set V consists of transactions. We seek to generate an association graph on the items such that only statistically significant associations are present.  
<br>
Consider the $N\times K$incidence matrix B that represents our data:  
<br>

$B=\left( \begin{matrix}
   1(\text{item 1}\in \text{transaction 1}) & \cdots  & 1(\text{item 1}\in \text{transaction K})  \\
   \vdots  & \ddots  & \vdots   \\
   1(\text{item N}\in \text{transaction 1}) & \cdots  & 1(\text{item N}\in \text{transaction K})  \\
\end{matrix} \right)$ 

<br>
Now define a new variable ${{x}_{ij}}$ that we can refer to as the “opportunities,” which is the number of transaction that either item I or item j link to. This can be obtained by performing an element-wise logical disjunction on the rows of B and summing the elements of the resulting vector: ${{x}_{ij}}=\sum\limits_{k=1}^{K}{OR({{b}_{ik,}}{{b}_{jk}})}$.  

<br>
From the B, we can obtain the co-occurrence matrix via 
<br>

$W=B{{B}^{T}}=\left( \begin{matrix}
   \sum\limits_{k=1}^{K}{1(\text{item 1}\in \text{transaction k})} & \cdots  & \sum\limits_{k=1}^{K}{1(\text{item 1 }\cap \text{ item N}\in \text{transaction k})}  \\
   \vdots  & \ddots  & \vdots   \\
   \sum\limits_{k=1}^{K}{1(\text{item N }\cap \text{ item 1}\in \text{transaction k})} & \cdots  & \sum\limits_{k=1}^{K}{1(\text{item N}\in \text{transaction k})}  \\
\end{matrix} \right)$ 
<br>

We can model each ${{w}_{ij}}\sim \text{binomial(}{{x}_{ij}},{{\pi }_{ij}}\text{)}$. We aim to infer the parameters ${{\pi }_{ij}}$ for all pairs $i,j$.  
<br>
We take a Bayesian approach and place a Beta distribution prior for each parameter. That is, ${{\pi }_{ij}}\sim \text{beta(}{{\alpha }_{ij}},{{\beta }_{ij}}\text{)}$.  Thus, can easily obtain the posterior as: $P({{\pi }_{ij}}|{{w}_{ij}},{{x}_{ij}})=\text{beta(}{{\alpha }_{ij}}+{{w}_{ij}},{{\beta }_{ij}}+{{x}_{ij}}-{{w}_{ij}}\text{)}$
<br>

#### Test

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(knitr)

# generate item pmf
nItems = 20
nSets = 5
prandom = 0.1

set.seed(3)
lengths = rpois(nSets, lambda = 2) + 1 # length of the 5 itemsets 

itemsets = vector("list", nSets)
for (i in seq_along(itemsets)) {itemsets[[i]]=sample(nItems,lengths[i])}
P_z = runif(nSets)
P_z = c((P_z/sum(P_z)) * (1-prandom), prandom)

### manipuate to render 
items = c(sapply(itemsets, function(x) paste("item", x,collapse=" ")) , "random")
item_pmf = as.data.frame(cbind(items,round(P_z,2)))
names(item_pmf) = c("Z = itemsets", "P(Z=z)")
kable(item_pmf)

```



```{r echo=TRUE, message=FALSE, warning=FALSE}
# simulate transactions from itemset PMF
nTrans = 200

B = array(0,dim=c(nTrans,nItems))

for (i in 1:nTrans) {
  j = rpois(1,1)+1 # how many itemsets does the transaction contain (add one to avoid empty transactions)
  k = findInterval(runif(j), cumsum(P_z)) + 1 # which itemsets
  if((nSets+1) %in% k) {itemsets[nSets+1] = sample(nItems,rpois(1,1)+1)} # if trans contains a random set, generate it
  B[i,unlist(itemsets[k])] = 1
}

B = t(B)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# generate 'PI' upper triangular matrix of attractions 
PI = array(0, dim = c(nItems, nItems))
alpha = 1;
beta = 1;

for (i in 1:(nItems-1)) {
  for (j in (i+1):nItems) {
    den = sum(ifelse(B[i,]+B[j,] > 0, 1, 0))
    if (den == 0) {PI[i,j] = 0
    } else {PI[i,j] = (B[i,] %*% B[j,]) / sum(ifelse(B[i,]+B[j,] > 0, 1, 0))}
  }
}
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# generate a graph from attraction matrix
library(igraph)

threshold = 0.4

edges <- c()
for (i in 1:(nItems-1)) {
  for (j in (i+1):nItems) {
    if (PI[i,j] > threshold)
      edges <- c(edges, i, j)
  }
}
g <- make_empty_graph(nItems, directed = FALSE) %>% add_edges(edges)
plot(g)

```



